Sparsity Weak Lensing $3$-D Density Map Reconstruction: Reducing the line of sight smearing.

### 2020-10-29 doc/paper_ms_method_HSCY1/ms.tex
update the draft of paper

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak-lensing shear measurements.
The $3$-D density contrast maps are modeled as a summation of the NFW basis atoms which have $2$-D multi-scale NFW surface
density profiles on the transverse plane and $1$-D Dirac delta functions in the line of sight direction. With the prior
assumption that the density fields can be sparsely represented by the NFW basis atoms, the density fields are reconstructed 
using an oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape noise and
photo-$z$ uncertainty.
Our findings are summarized as folllows: 1) The global minima solution of the standard lasso algorithm suffers from a
smear of structure in the line of sight direction even in the absence of shape nosie.
2) The global minima solution of addaptive lasso algorithm significantly removes the line of sight smear of structure.
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis atoms in Fourier space
            and the lower row shows the basis atoms in Real space.
            The leftmost column is point mass atom and the other columns are multi-scale NFW atoms.
            The smoothing kernel is Gaussian with standard deviation of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

@@

The expected shear measurements ($\gamma$) on distant galaxies are related to the
foreground density contrast field ($\delta$) through a linear transformation
\begin{equation} \label{eq-intro-delta2shear}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is used to denote the linear transformation operator which includes not only physical lensing effect
but also systematic effects from observations (e.g. pixelization and smoothing of shear field in
transverse plane, photo-$z$ uncertainty).

@@

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection parameters to the density contrast,
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

@@

$N$-body simulations have shown that the dark matter is distributed in halos connected by filaments, and the density
profile of a single halo follows the NFW function \citep{halo-NFW1997ApJ}.
We construct a dictionary with multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles of NFW
functions \citep{haloModel-TJ2003-3pt} on the transverse plane. We follow \citet{LSS-massMap-Glimpse3D-Leonard2014}
to neglect the depth of halos since the resolution scale of the reconstruction in the line of sight direction is much
larger than the scales of halos. Therefore, we set the profile of the NFW atoms in the line of sight as the Dirac delta
function. The multi-scale NFW atoms are used to model the profiles of halos which is assumed to be sparsely distributed
in the universe. With the sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is used to reconstruct
the density field. We find that, in contrast to the standard lasso estimator which smears the structure along the line of sight,
the addaptive lasso is able to significantly reduce the smearing effect.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their MLZ best photo-$z$ estimation.
        The blue histogram is the number distribution of the best photo-$z$ estimation. The vertical dashed lines
        are the boundaries of the bins. The galaxies are equal-number binned.} \label{fig-bestpz}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includezraphics{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includezraphics{delta-3-7-pz-nn-alasso.pdf}}
\caption{The result of density map reconstruction with standard lasso (left) and adaptive lasso (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction
        smear the density field along the line of sight direction (vertical direction in the plots). Noises on shear measurements
        are neglected in this simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

The density contrast field is modeled as a summation of basis atoms in the dictionary
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms have `$N$' different scale frames and the atoms
in each scale frame are shifted by $\vec{r'}$ to form basis at different positions. $x_s(\vec{r'})$ is the projection
coefficients of the density contrast field onto the basis atoms.

We propose to use multi-scale NFW atoms, denoted as $\{\phi_1,...,\phi_N\}$, as our dictionary basi.
On the transverse plane, the NFW atoms follow surface density profiles of NFW halos
\citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and truncation radius $c \theta_\alpha$ ,
where $c$ is the concentration of NFW halos.
As the scales of halos are much less than the reachable redshift resolution, we neglect the depth of halo on the
line of sight direction and set the profiles of NFW atoms in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}.
The multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for NFW atoms in different scale frames.

To simplify the notation, we define the projection parameters as a column vector:
$x=\begin{pmatrix}
x_{0}\\
x_{1}\\
...\\
x_{N}
\end{pmatrix}$,
and the dictionary transform operator as a row vector:
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
We substitute eq. (\ref{eq-x2delta}) into eq.(\ref{eq-delta2gammat})
\begin{equation}\label{eq-x2gammat}
\gamma_L=\mathbf{Q}\mathbf{\Phi} x.
\end{equation}

@@

The photometric redshifts of source galaxies in the current large scale survey are estimated with a limited number of
board photometric bands (e.g. $9$ bands for KIDS$+$VIKING survey \citep{KIDS_VIKING-Hildebrant2020}, $5$ bands for DES
survey and HSC survey). As a result, the estimated redshifts of galaxies suffer from much larger uncertainties, compared
with redshifts estimated with spectroscopic observations. Such photo-$z$ uncertainty smears the lensing kernels statistically
since a galaxy with a best fit photo-$z$ estimation of $z_s$ has possibilities of being actually located at different
redshifts ($z$). The probability function for the photo-$z$ uncertainty is denoted as $P(z|z_s)$ and the expected shear
distortion on the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainty changes the shear as
\begin{equation}
\gamma_L \rightarrow \mathbf{Q} \gamma_L,
\end{equation}
where we use $\rightarrow$ to denote the changes of shear by the systematic operator.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{shapeMeasurementError-HSCY1.pdf}
 \caption{The histograms of the HSC-like shear measurement error (including both from shape noise and from photon noise) 
        on the first component of shear ($g_1$) for galaxies (blue lines) and smoothed pixels (orange lines). The solid
        steps are the sampled histograms, and the dashed lines are the normalized gaussian distributions which have the
        same standard deviations as the corresponding histograms.}
 \label{fig:mass-redshift}
\end{figure}

@@

The smoothing is conducted by convolving the shear measurements with a smoothing kernel
\begin{equation}
\gamma_{\rm{sm}} (\vec{\theta})  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) \gamma_i}{\sum_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $\gamma_i$, $z_i$ and $\theta_i$ are the shear,
photometric reshift, and transverse position of the `$i$-th' galaxy in the catalog.

$W(\vec{\theta},z)$ can be decomposed into a transverse component $W_T(\vec{\theta})$ and a line of sight component
$W_\times(z)$
\begin{equation}
W(\vec{\theta},z)=W_T(\vec{\theta}) W_\times (z).
\end{equation}
In this paper, we use an isotropic $2$-D Gaussian kernel and a $1$-D top-hat kernel to smooth the measurements in the
transverse plane and the line of sight direction, the smoothing kernel is
\begin{equation}
\begin{split}
W_T(\vec{\theta}) &=\frac{1}{2\pi\beta^2}\exp(-\frac{|\vec{\theta}|}{2\beta^2}),\\
W_\times (z) &=
\begin{cases}
1/\Delta z& (|z|<\Delta z/2),\\
0& else.
\end{cases}
\end{split}
\end{equation}

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{noise_std_map_pix.pdf}
 \caption{The standard deviation pixel map of the HSC-like shear measurement error for the fifth source galaxy bin 
        ($0.69 \leq z < 0.80 $).}
\end{figure}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-3.pdf}
 \caption{The upper panels show the halo detection rate and the number of false peaks per square degree for mass map 
 reconstruction with NFW atoms and $\lambda=3.5$. The low panels show the results for $\lambda=5.0$. }
\end{figure*}

@@

We define the masking window function according to the smoothed number density (defined in eq. (\ref{eq-numDesSmooth})) of
the galaxies as follows
\begin{equation}
 M(\vec{r})=
\begin{cases}
0 & n_{\rm{sm}}>1,\\
1 & \rm{else}.
\end{cases}
\end{equation}
The mask changes the shear measurements though
\begin{equation}\label{eq-delta2gamma-final}
\gamma_L(\vec{\theta},z) \rightarrow M(\vec{\theta},z) \gamma_L(\vec{\theta},z),
\end{equation}
We define the masking operator as
\begin{equation}
\mathbf{M}= \int d^3 r' M(\vec{r'}) \delta_D(\vec{r}-\vec{r'}),
\end{equation}
where $\delta_D(\vec{r})$ is $3$-D Dirac delta function. The shear is influenced by the masking through
$\gamma_L \rightarrow \mathbf{M} \gamma_L$.

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{peak_scatters_f3-1.pdf}
 \caption{The left panel shows the stacked $2$-D distribution of the deviations of detected peak positions
 from the centers of the corresponding input halos. The $x$-axis is for the deviated distance in the transverse plane, and 
 the $y$-axis is for the deviation of the redshift. For each simulatin, the peaks inside the dashed black box with the
 minimal deviations (in unit of pixel) are defined as true detections. 
 The right panel focuses on the deviation of detected peaks in the line of sight direction. The $x$-axis is the input 
 halo redshifts, and the $y$-axis is the redshift of the detected peak. The `$\cross$' denotes the average redshift of 
 detected peaks for each halo over different realizations, and the error-bars are the scatters of the corresponding peak 
 redshifts. All of the results in this figure are based on the NFW dictionary with $\lambda=3.5$.}
\end{figure*}

We pixelize the smoothed shear field into a $N_\theta \times N_\theta \times N_s$ grid, where $N_\theta$
is the number of pixels of two orthogonal axes of the transverse plane and $N_s$ is the number of pixels of the
line of sight axis. $\gamma_{\alpha}$ is used to denote the value recorded on pixel with index $\alpha$, where $\alpha=
1...N_\theta \times N_\theta \times N_s$. The grids on the transverse plans are equally spaced, therefore fast Fourier
transform (FFT) can be used to boost the speed of linear operation on the transverse plan, whereas the grids in the
line of sight direction follow equal number binning as shown in Figure \ref{fig-bestpz}.

Similarly, we pixelize each scale frame of the parameter $x$ into a $N_\theta \times N_\theta \times N_l$ grid.
The pixelization of the parameter field $x$ on the transverse plan for each scale frame is exactly the same as the
pixelization of the smoothed shear field on the transverse plane. Whereas the parameter field is pixelized into equal
spaced grids in the line of sight direction. Here, we use $N_l$ to denote the number of the lens planes and $x_{\beta}$ to
denote the element of the parameter with index $\beta$, where $\beta=1...N_\theta \times N_\theta \times N_l \times N$.
The corresponding pixelized elements of the forward transform matrix $\mathbf{A}$ is denoted as $A_{\alpha\beta}$.

We term the column vectors of the forward transform matrix $\mathbf{A}$ as the effective basis atoms. We note that the
effective basis atoms have different $l^2$ norm. The $l^2$ norm of the $i'th$ column vectors of the effective basis atoms are
$\mathcal{N}_{i}=\sum_\alpha A_{i\alpha}A_{i\alpha}$. Before solving the density map reconstruction problem, we normalize
the column
vectors of the transform matrix through the following rescaling
\begin{equation}
\begin{split}
A'_{\alpha\beta}&=A_{\alpha\beta}/\mathcal{N}_{\alpha}^{\frac{1}{2}},\\
x'_{\beta}&=x_{\beta}\mathcal{N}_{\beta}^{\frac{1}{2}}.\\
\end{split}
\end{equation}

@@

The lasso algorithm uses $l^1$ norm of parameters as the regularization term and the estimator is defined as
\begin{equation}
\hat{x'}^{\rm{lasso}}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{A'} x')}_2^2+ \lambda_{\rm{ls}} \norm{x'}^1_1\right\},
\end{equation}
where $\norm{\bigcdot}_1$ and $\norm{\bigcdot}_2$ refer to the $l^1$ norm and $l^2$ norm, respectively.
$\lambda_{ls}$ denotes the penalization parameter for the lasso estimation.

The lasso algorithm selects the parameters which are relevant to the measurements and simultaneously estimate the value of
the selected parameters. However, it has been shown by
\citet{AdaLASSO-Zou2006} that when the column vectors of the projection matrix $\mathbf{A'}$ are highly correlated,
lasso cannot select the relevant parameters from the parameter space consistently. Moreover, the estimated parameters
are biased due to the shrinkage process in the lasso regression. We note that, for the density map reconstruction problem,
the effective basis vectors are highly correlated since, as shown in Figure \ref{fig-corlensKer}, the lensing kernels for
lenses at different redshifts are highly correlated. Therefore, the lasso algorithm cannot select the consistent mass in
the line of sight direction and the reconstructed mass sufferes from smears in the line of sight direction even in the
absence of noise on shear measurements.
Figure \ref{fig-lassoVsadaLasso} shows the reconstructions of the mass map of a single halo with halo mass equals
$M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The noises on shear measurements are not included. The left panel of
Figure \ref{fig-lassoVsadaLasso} is the reconstruction with lasso algorithm which shows a significant smear along the line of sight.

>>>>

\begin{abstract}
A new method is developed to reconstruct $3$-D density contrast maps from photometric weak-lensing shear measurements.
The $3$-D density contrast maps are modeled as a summation of the NFW basis atoms which have $2$-D multi-scale NFW surface
density profiles on the transverse plane and $1$-D Dirac delta functions in the line of sight direction. With the prior
assumption that the density fields can be sparsely represented by the NFW basis atoms, the density fields are reconstructed 
using an oracle algorithm: adaptive lasso. Our method is tested with realistic simulations using the HSC-like shape estimation
error and photo-$z$ uncertainty.
Our findings are summarized as folllows: 
1) The solution of the standard lasso algorithm suffers from a smear of structure 
in the line of sight direction even in the absence of shape nosie, and the addaptive lasso algorithm significantly removes 
the line of sight smear of structure. 
2) The algorithm is able to detect halo with minimal mass limits of $10^{13.5} M_{\odot}/h$, $10^{14.3} M_{\odot}/h$, 
$10^{15.0} M_{\odot}/h$ for the low ($z<0.3$), median ($0.3\leq z< 0.6$) and high ($0.6\leq z< 0.9$) redshifts, respectively.
3) The estimated redshift of the halos detected from the reconstructed mass maps are
lower than the true redshfit by about $0.03$ for halos at low redshifts ($z\leq 0.4$). The relative redshfit bias is below 
$0.5\%$ for halos at $0.4<z\leq 0.85$.
\end{abstract}

@@

\begin{figure*}[!t]
    \includegraphics[width=1.\textwidth]{nfwlet-atom-2D.pdf}
    \caption{The smoothed pixelized basis atoms. The upper row shows the basis atoms in Fourier space
            and the lower row shows the basis atoms in Real space.
            The leftmost column is the point mass atom and the other columns are the multi-scale NFW atoms.
            The smoothing kernel is Gaussian with standard deviation of $1.5$ pixels.} \label{fig-atoms2D}
\end{figure*}

@@

The expected shear measurements ($\gamma$) on distant galaxies are related to the
foreground density contrast field ($\delta$) through a linear transformation:
\begin{equation} \label{eq-intro-delta2shear}
 \gamma=\mathbf{T} \delta,
\end{equation}
where $\mathbf{T}$ is used to denote the linear transformation operator which includes not only physical lensing effect
but also systematic effects from observations (e.g. pixelization and smoothing of shear field in
transverse plane, photo-$z$ uncertainty).

@@

In order to fully reconstruct the $3$-D mass density distribution ($\delta$) from the photometric shear observations ($\gamma$),
the density contrast field is modeled as a summation of basis atoms in a model dictionary:
\begin{equation} \label{eq-intro-dict}
 \delta= \mathbf{\Phi} x,
\end{equation}
where $\mathbf{\Phi}$ is the transformation operator from the projection coefficient field to the density contrast,
and $x$ denotes the parameters. \citet{LSS-massMap-Wiener-Simon2009} reconstruct the density field
in Fourier space, which is equivalent to model the mass field with sinusoidial functions. On the other hand,
\citet{LSS-massMap-Glimpse3D-Leonard2014} models the mass field with Starlets \citep{Starlet-Starck2015}.

@@

$N$-body simulations have shown that the dark matter is distributed in halos connected by filaments, and the density
profile of a single halo follows the NFW function \citep{halo-NFW1997ApJ}.
We construct a dictionary with the multi-scale NFW atoms.  The atoms follow multi-scale surface density profiles of the 
NFW functions \citep{haloModel-TJ2003-3pt} on the transverse plane. We follow \citet{LSS-massMap-Glimpse3D-Leonard2014}
to neglect the depth of halos since the resolution scale of the reconstruction in the line of sight direction is much
larger than the scales of halos. Therefore, we set the profile of the NFW atoms in the line of sight as the Dirac delta
function. The multi-scale NFW atoms are used to model the profiles of halos which is assumed to be sparsely distributed
in the universe. With the sparsity prior, the adaptive lasso regularization \citep{AdaLASSO-Zou2006} is used to reconstruct
the density field. We find that, in contrast to the standard lasso estimator which smears the structure along the line of sight,
the addaptive lasso is able to significantly reduce the smearing effect.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.5\textwidth]{photo-z_binning.pdf}
 \caption{The source galaxies are binned into $10$ redshift bins according to their Machine Learning and photo-Z (MLZ) 
        best photo-$z$ estimation. The blue histogram is the normalized number distribution of the best photo-$z$ estimation. 
        The vertical dashed lines are the boundaries of the bins. The galaxies are equal-number binned.} \label{fig-bestpz}
\end{figure}

@@

\begin{figure*}[!t]
\centering
\subfigure[lasso]{\includezraphics{delta-3-7-pz-nn-lasso.pdf}}
\subfigure[adaptive lasso]{\includezraphics{delta-3-7-pz-nn-alasso.pdf}}
\caption{The results of density map reconstructions with standard lasso (left) and adaptive lasso (right). The input density
        map is from a NFW halo with mass $M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The lasso reconstruction
        smear the density field along the line of sight direction (vertical direction in the plots). The vertical direction
        is the line of sight direction, and the lower boundaries and upper boundaries of the boxes are $0.01$ and $0.85$, 
        respectively. Noises on galaxy shape measurements are neglected in this simulation.} \label{fig-lassoVsadaLasso}
\end{figure*}

The density contrast field is modeled as a summation of basis atoms in the dictionary:
\begin{equation}\label{eq-x2delta}
\delta(\vec{r}) = \sum_{s=1}^{N} \int d^3 r' \phi_s(\vec{r}-\vec{r'}) x_s(\vec{r'}),
\end{equation}
where $\phi_s(\vec{r})$ are the basis atoms of the dictionary. The basis atoms have `$N$' different scale frames and the atoms
in each scale frame are shifted by $\vec{r'}$ to form basis at different positions. $x_s(\vec{r'})$ is the projection
coefficients of the density contrast field onto the basis atoms.

We propose to use the multi-scale NFW atoms, denoted as $\{\phi_1,...,\phi_N\}$, as the basis atoms of our dictionary .
On the transverse plane, the NFW atoms follow surface density profiles of the NFW halos
\citep{haloModel-TJ2003-3pt} with scale radius $\theta_\alpha$ and truncation radius $c \theta_\alpha$ ,
where $c$ is the concentration of the NFW halos.
As the scales of halos are much less than the reachable redshift resolution, we neglect the depth of halo on the
line of sight direction and set the profiles of the NFW atoms in the line of sight direction to $1$-D Dirac delta
functions as suggested by \citep{LSS-massMap-Glimpse3D-Leonard2014}.
The multi-scale NFW atoms are defined as
\begin{equation}
\begin{split}
\phi_\alpha(\vec{r}) =&\frac{f }{2 \pi \theta_\alpha^2 } F(|\vec{\theta}|/\theta_\alpha) \delta_D(z),\\
&  (s=1..N)
\end{split}
\end{equation}
where
\begin{equation}
F(x)=
\begin{cases}
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccosh \left(\frac{x^2+c}{x(1+c)}\right)}{(1-x^2)^{3/2}}  & (x<1),\\
\frac{\sqrt{c^2-1}}{3(1+c)} (1+\frac{1}{c+1}) & (x=1),\\
-\frac{\sqrt{c^2-x^2}}{(1-x^2)(1+c)} + \frac{\arccos\left(\frac{x^2+c}{x(1+c)}\right)}{(x^2-1)^{3/2}} & (1<x\leq c),\\
0& (x>c).
\end{cases}
\end{equation}
$f=1/[\ln (1+c)-c/(1+c)]$. In this work, we fix $c=4$ for the NFW atoms in different scale frames.

To simplify the notation, we define the projection coefficients as a column vector:
$x=\begin{pmatrix}
x_{0}\\
x_{1}\\
...\\
x_{N}
\end{pmatrix}$,
and the dictionary transform operator as a row vector:
\begin{equation}
\mathbf{\Phi}=\begin{pmatrix}
\int d^3r\phi_0(\vec{r}) ~\int d^3r \phi_1(\vec{r})~ ...~\int d^3r \phi_{N}(\vec{r})
\end{pmatrix}.
\end{equation}
We substitute eq. (\ref{eq-x2delta}) into eq.(\ref{eq-delta2gammat}) and get
\begin{equation}\label{eq-x2gammat}
\gamma_L=\mathbf{Q}\mathbf{\Phi} x.
\end{equation}

@@

The photometric redshifts of source galaxies in the current large scale survey are estimated with a limited number of
board photometric bands (e.g. $9$ bands for KIDS$+$VIKING survey \citep{KIDS_VIKING-Hildebrant2020}, $5$ bands for DES
survey and HSC survey). As a result, the estimated redshifts of galaxies suffer from much larger uncertainties, compared
with redshifts estimated with spectroscopic observations. Such photo-$z$ uncertainty smears the lensing kernels statistically
since a galaxy with a best fit photo-$z$ estimation of $z_s$ has possibilities of being actually located at different
redshifts ($z$). The probability function for the photo-$z$ uncertainty is denoted as $P(z|z_s)$ and the expected shear
distortion on the galaxy is
\begin{equation}\label{eq-delta2gamma-poz}
\int dz_s P(z|z_s) \gamma_L(\vec{\theta},z_s).
\end{equation}
With the definition of photo-$z$ smearing operator:
\begin{equation}
\mathbf{P} = \int dz_s P(z|z_s),
\end{equation}
the photo-$z$ uncertainty changes the shear as
\begin{equation}
\gamma_L \rightarrow \mathbf{Q} \gamma_L,
\end{equation}
where we use $\rightarrow$ to denote the changes of shear by the systematic operator.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{shapeMeasurementError-HSCY1.pdf}
 \caption{The histograms of the HSC-like shape measurement error (including both from shape noise and from photon noise) 
        on the first component of shear ($g_1$) for galaxies (blue lines) and smoothed pixels (orange lines). The solid
        steps are the sampled histograms, and the dashed lines are the normalized gaussian distributions which have the
        same standard deviations as the corresponding histograms.}
 \label{fig-noiseHistogram}
\end{figure}

@@

The smoothing is conducted by convolving the shear measurements with a smoothing kernel:
\begin{equation}
\gamma_{\rm{sm}} (\vec{\theta})  = \frac{\sum_i  W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) \gamma_i}{\sum_i W(\vec{\theta}-\vec{{\theta}}_i,z-z_i) },
\end{equation}
where $W(\vec{\theta},z)$ is a $3$-D smoothing kernel. $\gamma_i$, $z_i$ and $\theta_i$ are the shear,
photometric reshift, and transverse position of the `$i$-th' galaxy in the catalog.

$W(\vec{\theta},z)$ can be decomposed into a transverse component $W_T(\vec{\theta})$ and a line of sight component
$W_\times(z)$ as
\begin{equation}
W(\vec{\theta},z)=W_T(\vec{\theta}) W_\times (z).
\end{equation}
In this paper, we use an isotropic $2$-D Gaussian kernel and a $1$-D top-hat kernel to smooth the measurements in the
transverse plane and the line of sight direction, the smoothing kernel is
\begin{equation}
\begin{split}
W_T(\vec{\theta}) &=\frac{1}{2\pi\beta^2}\exp(-\frac{|\vec{\theta}|}{2\beta^2}),\\
W_\times (z) &=
\begin{cases}
1/\Delta z& (|z|<\Delta z/2),\\
0& else.
\end{cases}
\end{split}
\end{equation}
In this paper, we set $\beta=1.\arcmin5$.

@@

\begin{figure}[!t]
 \centering
 \includegraphics[width=0.45\textwidth]{noise_std_map_pix.pdf}
 \caption{The standard deviation pixel map of the HSC-like shape measurement error for the fifth source galaxy bin 
        ($0.69 \leq z < 0.80 $).
        } \label{fig-noistdmap}
\end{figure}

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{peak_scatters_f3-1.pdf}
 \caption{The left panel shows the stacked $2$-D distribution of the deviations of detected peak positions
 from the centers of the corresponding input halos. The $x$-axis is for the deviated distance in the transverse plane, and 
 the $y$-axis is for the deviation of the redshift. For each box, the positive peaks inside the dashed black box with 
 the minimal offset from the input halo (in unit of pixel) is taken as a true detection. 
 The right panel focuses on the deviation of detected peaks in the line of sight direction. The $x$-axis is the input 
 halo redshifts, and the $y$-axis is the redshift of the detected peak. The `$\cross$' denotes the average redshift of 
 detected peaks for each halo over different noise realizations, and the error-bars are the uncertainties of the average
 redshifts. The deep gray area is for the relative redshfit bias less than $0.05$, and the light gray area is for the 
 relative redshift bias less than $0.5$. These results in this figure are based on the NFW dictionary with $\lambda=3.5$.
 } \label{fig-detoffsets}
\end{figure*}

@@

We define the masking window function according to the smoothed number density (defined in eq. (\ref{eq-numDesSmooth})) of
the galaxies:
\begin{equation}
 M(\vec{r})=
\begin{cases}
0 & n_{\rm{sm}}>1,\\
1 & \rm{else}.
\end{cases}
\end{equation}
The mask changes the shear measurements though
\begin{equation}\label{eq-delta2gamma-final}
\gamma_L(\vec{\theta},z) \rightarrow M(\vec{\theta},z) \gamma_L(\vec{\theta},z),
\end{equation}
We define the masking operator as
\begin{equation}
\mathbf{M}= \int d^3 r' M(\vec{r'}) \delta_D(\vec{r}-\vec{r'}),
\end{equation}
where $\delta_D(\vec{r})$ is $3$-D Dirac delta function. The shear is influenced by the masking through
$\gamma_L \rightarrow \mathbf{M} \gamma_L$.

@@

\begin{figure*}[!t]
 \centering
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-1.pdf}
 \includegraphics[width=0.95\textwidth]{detfalseRate_f3-3.pdf}
 \caption{The upper panels show the halo detection rate and the number of false peaks per square degree for mass map 
        reconstructed with the NFW atoms. The penalization parameter is set to $\lambda=3.5$ for the upper panels. 
        The low panels show the results for $\lambda=5.0$. 
        } \label{fig-detFalsRateNFW}
\end{figure*}

We pixelize the smoothed shear field into a $N_\theta \times N_\theta \times N_s$ grid, where $N_\theta$
is the number of pixels for the two orthogonal axes of the transverse plane and $N_s$ is the number of pixels for the
line of sight axis. $\gamma_{\alpha}$ denotes the smoothed shear measurements recorded on the pixel with index $\alpha$, 
where $\alpha=1...N_\theta \times N_\theta \times N_s$. The grids on the transverse plans are equally spaced with a pixel
size of $1\arcmin$.
Fast Fourier transform (FFT) can be used to boost the speed of linear operation on the transverse plan, whereas the grids 
in the line of sight direction follow equal number binning as shown in Figure \ref{fig-bestpz}.

Similarly, we pixelize each scale frame of the projection coefficient field $x$ into a $N_\theta \times N_\theta \times N_l$ 
grid. The pixelization on the transverse plan for each scale frame is exactly the same as that of the smoothed shear field 
on the transverse plane. Whereas the projection coefficient field is pixelized into equal spaced grids in the line of sight 
direction ranging from redshfit $0.01$ to redshift $0.85$. Here, we use $N_l$ to denote the number of the lens planes and 
$x_{\beta}$ to denote the element of the projection parameter with index $\beta$, where 
$\beta=1...N_\theta \times N_\theta \times N_l \times N$. The corresponding pixelized elements of the forward transform 
matrix $\mathbf{A}$ is denoted as $A_{\alpha\beta}$.

We term the column vectors of the forward transform matrix $\mathbf{A}$ as the effective basis atoms. We note that the
effective basis atoms have different $l^2$ norm. The $l^2$ norm of the $i'th$ column vectors of the effective basis atoms are
$\mathcal{N}_{i}=\sum_\alpha A_{i\alpha}A_{i\alpha}$. Before solving the density map reconstruction problem, we normalize
the column
vectors of the transform matrix through a rescaling:
\begin{equation}
\begin{split}
A'_{\alpha\beta}&=A_{\alpha\beta}/\mathcal{N}_{\alpha}^{\frac{1}{2}},\\
x'_{\beta}&=x_{\beta}\mathcal{N}_{\beta}^{\frac{1}{2}}.\\
\end{split}
\end{equation}

@@

The standard lasso algorithm uses the $l^1$ norm of the projection coefficient field as the regularization and the estimator is 
defined as
\begin{equation}
\hat{x'}^{\rm{lasso}}=\argmin_{x} \left\{ \frac{1}{2}\norm{\Sigma^{-\frac{1}{2}}(\gamma- \mathbf{A'} x')}_2^2+ \lambda_{\rm{ls}} \norm{x'}^1_1\right\},
\end{equation}
where $\norm{\bigcdot}_1$ and $\norm{\bigcdot}_2$ refer to the $l^1$ norm and $l^2$ norm, respectively.
$\lambda_{ls}$ denotes the penalization parameter for the lasso estimation.

The lasso algorithm selects the parameters which are relevant to the measurements and simultaneously estimate the value of
the selected parameters. However, it has been shown by
\citet{AdaLASSO-Zou2006} that when the column vectors of the transforming matrix $\mathbf{A'}$ are highly correlated,
lasso cannot select the relevant parameters from the parameter space consistently. Moreover, the estimated parameters
are biased due to the shrinkage process in the lasso regression. We note that, for the density map reconstruction problem,
the column vectors are highly correlated because, as shown in Figure \ref{fig-corlensKer}, the lensing kernels for
lenses at different redshifts are highly correlated. Therefore, the lasso algorithm cannot select the consistent mass in
the line of sight direction and the reconstructed mass sufferes from smears in the line of sight direction even in the
absence of noise on shear measurements.
Figure \ref{fig-lassoVsadaLasso} shows the reconstructions of the mass map of a single halo with halo mass equals
$M_{200}=10^{15} ~h^{-1}M_{\odot}$ at redshift $0.35$. The noises on shear measurements are not included in the simulation. 
The left panel of Figure \ref{fig-lassoVsadaLasso} is the reconstruction with lasso algorithm which shows a significant smear 
along the line of sight.
